I would like to thank my supervisor, James Danckert, for all of the support and
encouragement over the many years that I have worked with him. I began work in
his lab by writing a few lines of code for his students 10
years ago, and through the many projects over the following
years, I believe that I have learned a great deal from him and am enthusiastic
about applying those lessons in whatever endeavours await in my future.
Committee members Britt Anderson and Mike Dixon also provided a great deal of
feedback over the duration of my time here, and I cannot
overstate the quality and usefulness of their suggestions and criticisms.

I would also like to thank Nadine Quehl for her tireless work maintaining the
neurological patient database and providing enthusiastic assistance whenever a
need arose. I am also greatly indebted to her predecessors, Danielle Striemer,
and Andrea Schneider for building that database. Of course, this research could
not have been completed without the countless hours volunteered by all those
who have participated. I cannot express enough gratitude for their help, and
hope they can be proud of their contribution to the progress of
science.

Next, I would like to thank my colleagues for the support and mentorship that
can only come from caring individuals with insatiably curious minds and an
unwavering ethic of camaraderie. In particular, Chris Striemer made my first
year of graduate school both productive and exciting, instilling many of the
core values that helped me develop as a researcher and complete these projects.
Marc Hurwitz, Grayden Solman, Linda Carson, and Alex Filipowicz provided a
fountain of inspiration, immeasurable support, encouragement, and an
inexhaustible supply of ideas. I am greatly indebted to these individuals, the
rest of the Danckert Attention and Action Group, and the whole of the Cognitive
Neuroscience division.

Lastly, there is one individual, without whom I could not have done any of
this. My wife, Annette Locklin, was my foundation and my motivation. I finished
this for us, because she was part of every struggle, every frustration, and
every step forward.
